---
title: "Lab 02 - Plastic waste"
author: "Lindsay Stall"
date: "01/20/2023"
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1.1

```{r histogram}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = .02) +
  labs(
    x = "Plastic Waste per Capita",
    title = "Plastic Waste Per Capita by Continent"
  ) +
facet_wrap(~ continent, nrow = 2) 

```


North America has the greatest variance in plastic waste per capita.


### Exercise 2.1

```{r plastic-waste-density}
ggplot(data = plastic_waste, 
       mapping = aes(x = plastic_waste_per_cap, 
                     color = continent, 
                     fill = continent)) +
  geom_density(alpha = 0.2)

```


Exercise 2.2: We defined the color and fill of the curves by mapping aesthetics of the plot because they are a function of the variable, continent. We defined the alpha level, however, as a characteristic of the plotting geom because alpha level does not change a function of the value of a variable (e.g., all continents have the same alpha leveL).


### Exercise 3.1

```{r box}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_boxplot()
```
```{r violin}
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
geom_violin()
```

Unlike a box plot, the violin plot presents a distribution of data such that we can actually see the shape of the data. Box plots are more simplified, but do provide some information that violin plots do not: for example, the dots on the box plot represent outliers, whereas there are not dots to represent outliers in violin plots.



### Exercise 4.1 - 4.3

```{r Exercise 4.1}
ggplot(data = plastic_waste, aes(x = mismanaged_plastic_waste_per_cap, y = plastic_waste_per_cap)) +
  geom_point()
```
```{r Exercise 4.2}
ggplot(data = plastic_waste, 
       aes(x = mismanaged_plastic_waste_per_cap, 
           y = plastic_waste_per_cap,
       color = continent)) +
  geom_point()

```
```{r Exercise 4.3}
ggplot(data = plastic_waste, 
       aes(x = total_pop, 
           y = plastic_waste_per_cap)) +
  geom_point()
```


```{r Exercise 4.3 pt 2}
ggplot(data = plastic_waste, 
       aes(x = coastal_pop, 
           y = plastic_waste_per_cap)) +
  geom_point()
```


Plastic waste per capita and coastal population appear to be more strongly linearly associated.


### Exercise 5.1

```{r}

df<- filter(plastic_waste, plastic_waste_per_cap < 3)
```


```{r recreate-viz}
ggplot(data = df, 
       aes(x = coastal_pop/total_pop, 
        y = plastic_waste_per_cap)) +
  labs(
    x = "Costal Population Proportion",
    y = "Plastic Waste per Capita",
    title = "Plastic Waste vs Costal Population Proportion"
  ) +
  geom_smooth()+
  geom_point(aes(color = continent))
```



